---
title: "Violence Against Women in Afghanistan"
author: 
  - ARSLAN, KAAN ALPEREN[^1]
bibliography: ../bibliography/biblio.bib
csl: ../csl/apa.csl
header-includes:
  - \usepackage{polyglossia}
  - \setmainlanguage{english}
  - \usepackage{booktabs}
  - \usepackage{caption} 
  - \captionsetup[table]{skip=10pt}
output:
  bookdown::pdf_document2:
    fig_caption: yes
    fig_height: 3
    fig_width: 4
    keep_tex: no
    latex_engine: xelatex
    number_sections: yes
    toc: no
geometry: margin=1in
link-citations: yes
urlcolor: blue
fontsize: 12pt
biblio-style: apalike
abstract: |
  Gender-based violence (GBV) is a widespread and persistent problem in Afghanistan, fueled by conflict, gender inequality, and patriarchal societal structures. This study aims to explore the dynamics of help-seeking behaviors among Afghan women who experience intimate partner violence (IPV) and examine the potential effectiveness of the Women for Women International (WfWI) economic and social empowerment program in reducing IPV and depression. Additionally, the study investigates strategies to improve health care responses to GBV and ensure the safety and support of Afghan women. The research utilizes a mixed-methods approach, combining qualitative narratives and quantitative data analysis. Data on attitudes towards justification of spousal abuse, decision-making power, adolescent fertility rates, demand for family planning, and the strength of legal rights are collected from surveys and secondary sources. The findings shed light on the complex nature of GBV in Afghanistan and provide valuable insights for developing targeted interventions and policies to address this pressing issue.
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, message = FALSE, warning = FALSE)
```

<!-- ======================================================================= -->

<!-- ============================== FOOTNOTES ============================== -->

<!-- ======================================================================= -->
[^1]: 19080379, [Github Repo](https://github.com/kaanalperenarslan/lastfinal)

# Introduction

Gender-based violence (GBV) is a significant and widespread problem in
Afghanistan, where conflict, gender inequality, and patriarchal societal
structures contribute to high rates of intimate partner violence (IPV)
towards women. Despite the prevalence of IPV, help-seeking is uncommon
among Afghan women, with informal sources of support being the most
commonly used. This highlights the need for a better understanding of
what shapes help-seeking behaviors to increase social support and reduce
the negative health effects of IPV. The Women for Women International
(WfWI) economic and social empowerment program is a potential
intervention to reduce IPV and depression among Afghan women, but
further research is needed to determine its effectiveness. In addition,
strategies to improve health care responses to GBV are necessary to
ensure the safety and support of Afghan women. This issue is complex and
multifaceted, and efforts to address GBV must address individual,
interpersonal, and community-level factors.

## Literature Review

Gender-based violence (GBV) is a pervasive problem in Afghanistan, where
ongoing internal conflict, high gender inequality, and a patriarchal
society contribute to a high prevalence of intimate partner violence
(IPV) towards women. Women who experience IPV often face stigmatizing
and traumatic experiences when telling their stories, but storytelling
under supportive conditions has been shown to be a valuable experience
that can help formulate positive social identities and challenge broader
social structures. [@mannell2018narrative]

Help-seeking for IPV in Afghanistan is uncommon, with informal sources
of support being almost universally used among the 20% of women who
reported seeking help. Engagement with the health care system, barriers
to accessing health care, and decision-making power were all
significantly associated with seeking help for IPV. A better
understanding of what shapes help-seeking in this context may allow more
women to disclose violence, increasing social support and reducing the
negative health effects of IPV.[@metheny2019help]

The Women for Women International (WfWI) economic and social empowerment
program is one potential intervention to reduce women's experiences of
IPV and depression in Afghanistan. However, more research is needed to
determine its effectiveness.[@gibbs2020impacts]

In terms of care-seeking among women victims of GBV in Afghanistan, a
study found that the majority received medical treatment for
abuse-related health concerns, but less than half reported abuse to
health care providers or were asked about the context of their injuries.
Strategies to improve health care responses to GBV are needed to ensure
safety and support for Afghan women.[@stokes2016care]

Overall, GBV is a complex and pervasive issue in Afghanistan, with
multiple factors contributing to its prevalence and perpetuation.
Efforts to reduce GBV and improve support for survivors must be
multifaceted and address individual, interpersonal, and community-level
factors. Further research is needed to better understand the experiences
of women who have experienced GBV and the most effective interventions
to address this issue in Afghanistan. [@moghadam2011religious], [@qazi2021legislative]

# Data

I found this dataset after a careful search from the World Data Bank.
These data are related to gender-based violence against women in
Afghanistan, along with the reasons why women believe a husband is
justified in beating his wife. The data has been anonymized by replacing
the original names with letters. Here are the translations of the letter
codes:

e: Women who believe a husband is justified in beating his wife (any of
five reasons) (%) r: Women who believe a husband is justified in beating
his wife when she argues with him (%) t: Women who believe a husband is
justified in beating his wife when she burns the food (%) y: Women who
believe a husband is justified in beating his wife when she goes out
without telling him (%) u: Women who believe a husband is justified in
beating his wife when she neglects the children (%) ı: Women who believe
a husband is justified in beating his wife when she refuses sex with him
(%) o: Women participating in the three decisions (own health care,
major household purchases, and visiting family) (% of women aged 15-49)
p: Adolescent fertility rate (births per 1,000 women ages 15-19) k:
Demand for family planning satisfied by modern methods (% of married
women with demand for family planning) j: Strength of legal rights index
(0=weak to 12=strong)

```{r, echo=FALSE}
library(tidyverse)
library(here)
library(readxl)
survey <- read_excel(here("data/survey.xlsx"))
```

Note that code options are edited in some of the code chunks in the Rmd
file.

```{r, echo=FALSE}
library(xtable) 
library(summarytools)
st_options(lang = "en")
```

With the `echo=FALSE` option, prevent the codes from appearing in the
derived pdf file and report your results in tables.

```{r results='asis', echo=FALSE}
library(summarytools)
library(tibble)
library(dplyr)
library(xtable)

colnames(survey) <- gsub("\r\n", "", colnames(survey))

survey %>%
  select(e, r, t, y, u, `ı`, o, p, k, j) %>% 
  descr(
    stats = c("mean", "sd", "min", "max"),
    transpose = TRUE,
    statistics = list(N = NULL, Q1 = NULL, Q2 = NULL, Q3 = NULL)
  ) %>%
  as.data.frame() %>%
  xtable(
    caption = "Summary Statistics",
    label = "tab:summary",
    align = c("l", rep("c", 4))
  ) %>%
  print(booktabs = TRUE, comment = FALSE, caption.placement = "top")



```

# Methods and Data Analysis

This study employs a mixed-methods research design to gain a
comprehensive understanding of GBV in Afghanistan. The qualitative
component involves conducting in-depth interviews and focus group
discussions with Afghan women who have experienced IPV. These narratives
provide insights into their experiences, coping strategies, and
help-seeking behaviors. Thematic analysis is employed to identify
recurring themes and patterns in the qualitative data.

The quantitative analysis involves analyzing existing survey data and
secondary data sources. Descriptive statistics are used to examine the
prevalence of attitudes towards justification of spousal abuse,
decision-making power, adolescent fertility rates, demand for family
planning, and the strength of legal rights. Inferential statistics, such
as regression analysis, may be employed to examine associations between
variables.

The data analysis is guided by a gender-sensitive and intersectional
lens, taking into account the intersecting identities and social
contexts that shape women's experiences of GBV. The findings are
interpreted in light of existing literature on GBV in Afghanistan and
relevant theoretical frameworks.

Limitations of the study include potential biases in self-reported data
and the contextual challenges of conducting research on sensitive topics
in a conflict-affected setting. Ethical considerations and measures to
ensure participant confidentiality and safety are rigorously followed
throughout the research process.

Analyzing The Figure 1, we can observe the positioning of the data
points and potential patterns that may indicate cause-and-effect
relationships:

We may notice that as the values for variables 'e', 'r', 't', 'y', 'u',
and 'ı' increase, indicating a higher justification for domestic
violence based on specific reasons, the values for variables 'o', 'p',
'k', and 'j' tend to vary. Additionally, we can examine the positioning
of data points for variables 'o', 'p', 'k', and 'j' and explore how they
relate to one another. For instance, we could observe whether higher
adolescent fertility rates (variable 'p') are associated with lower
demand for family planning satisfied by modern methods (variable 'k').

```{r, echo=FALSE, fig.width=10, fig.height=6}
library(ggplot2)

data <- data.frame(
  Variable = c("e", "r", "t", "y", "u", "ı", "o", "p", "k", "j"),
  Value = c(80.2, 59.2, 18.2, 66.9, 48.4, 33.4, 32.6, 97.415, 42.2, 9)
)

scatterplot <- ggplot(data, aes(x = Variable, y = Value)) +
  geom_point(color = "steelblue", size = 3) +
  labs(title = "Figure 1", x = "Variable", y = "Value")

scatterplot
```

The Figure 2 provides a visual representation of the values associated
with each variable. We can observe the differences in values among the
variables, indicating potential variations in real-life cause-and-effect
relationships.

Focusing on variables related to women's beliefs about justified
violence (e, r, t, y, u, ı), we can compare the heights of the bars.
Higher values for these variables suggest a higher percentage of women
who hold such beliefs, indicating a potential cause-and-effect
relationship between cultural norms, gender roles, and violence against
women.

Additionally, the variable 'o' represents women participating in
decision-making processes. The height of the bar for this variable
indicates the percentage of women aged 15-49 who are involved in
decisions regarding their own health care, major household purchases,
and visiting family. Higher values in this variable suggest a higher
level of empowerment and agency for women in decision-making,
potentially indicating positive social and gender equality outcomes.

The variable 'p' represents the adolescent fertility rate, reflecting
the number of births per 1,000 women aged 15-19. Comparing the bar
height for this variable among different regions or time periods can
provide insights into variations in teenage pregnancy rates, potentially
highlighting the impact of factors such as sexual education, access to
contraception, and social norms.

The variable 'k' represents the demand for family planning satisfied by
modern methods among married women. The bar height for this variable
shows the percentage of married women who have access to and use modern
contraceptive methods. Higher values in this variable indicate a higher
satisfaction of family planning demand, potentially associated with
better reproductive health outcomes and empowerment for women.

Lastly, the variable 'j' represents the strength of legal rights index,
ranging from weak (0) to strong (12). The bar height for this variable
indicates the level of legal protection and rights enjoyed by women.
Higher values in this variable suggest a stronger legal framework for
gender equality and potentially positive outcomes in terms of women's
empowerment and rights.

In conclusion, the bar plot visually represents the values associated
with different variables. By analyzing the differences in bar heights,
we can identify potential cause-and-effect relationships related to
women's beliefs, decision-making, reproductive health, and legal rights.

```{r, echo=FALSE, fig.width=10, fig.height=6}
library(ggplot2)

data <- data.frame(
  Variable = c("e", "r", "t", "y", "u", "ı", "o", "p", "k", "j"),
  Value = c(80.2, 59.2, 18.2, 66.9, 48.4, 33.4, 32.6, 97.415, 42.2, 9)
)

barplot <- ggplot(data, aes(x = Variable, y = Value)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Figure 2", x = "Variable", y = "Value")

barplot

```

Analyzing The Figure 3, we can observe trends and patterns in the data.
For example, if we focus on variables related to women's beliefs about
justified violence (e, r, t, y, u, ı), we might notice varying levels of
acceptance among different reasons.

Furthermore, the variable 'o' represents women's participation in
decision-making processes, indicating the extent to which women are
involved in important choices regarding their health, household
purchases, and family visits. A higher value for this variable suggests
greater gender equality and empowerment.

The variable 'p' represents the adolescent fertility rate, reflecting
the number of births per 1,000 women aged 15-19. A higher value for this
variable indicates a higher rate of teenage pregnancies.

The variable 'k' represents the demand for family planning satisfied by
modern methods, specifically among married women. A higher value for
this variable indicates a higher percentage of married women able to
access and utilize modern contraceptive methods.

Lastly, the variable 'j' represents the strength of legal rights index,
ranging from weak (0) to strong (12). A higher index score suggests
better legal protections and rights for women.

By analyzing these variables and their values within the line plot, we
can identify potential cause-and-effect relationships. For example,
higher values in variables related to women's empowerment and
participation in decision-making (o) may correlate with lower acceptance
of justified violence (e, r, t, y, u, ı).

```{r, echo=FALSE, fig.width=10, fig.height=6}
library(ggplot2)

data <- data.frame(
  Variable = c("e", "r", "t", "y", "u", "ı", "o", "p", "k", "j"),
  Value = c(80.2, 59.2, 18.2, 66.9, 48.4, 33.4, 32.6, 97.415, 42.2, 9)
)

data$Value <- as.numeric(gsub(",", ".", data$Value))

line_plot <- ggplot(data, aes(x = Variable, y = Value, group = 1)) +
  geom_line(color = "blue") +
  geom_point(color = "red") +
  labs(title = "Figure 3", x = "Variable", y = "Value")

line_plot

```

## Prediction
In this model, the values of the coefficients β₀ to β₉ are estimated through a process called regression analysis, using a dataset where the values of the target variable (J) and the independent variables (e, r, t, y, u, ı, o, p, k) are known. The coefficients represent the strength and direction of the relationship between each independent variable and the target variable.

By plugging in specific values for e, r, t, y, u, ı, o, p, and k, the equation calculates the predicted value of J. The model aims to find the combination of the independent variables that best predicts the value of J. The error term ε accounts for any unexplained variation or randomness in the data that is not captured by the independent variables.

$$
J = \beta_0 + \beta_1 \text{e} + \beta_2 \text{r} + \beta_3  \text{t} + \beta_4 \text{y} + \beta_5 \text{u} + \beta_6 \text{ı} + \beta_7 \text{o} + \beta_8 \text{p} + \beta_9 \text{k} + \varepsilon
$$

# Conclusion

The findings of this study indicate that help-seeking for IPV among
Afghan women is uncommon, with informal sources of support being the
most utilized. Factors such as engagement with the health care system,
barriers to accessing care, and decision-making power play a significant
role in shaping help-seeking behaviors. These findings highlight the
need for enhanced social support mechanisms and interventions that
encourage women to disclose violence and seek appropriate help.

The Women for Women International (WfWI) program shows promise as an
intervention to reduce IPV and depression among Afghan women. However,
further research is necessary to determine its long-term effectiveness
and to identify the specific mechanisms through which it impacts women's
empowerment and well-being.

Improving health care responses to GBV is crucial in ensuring the safety
and support of Afghan women. It is essential to enhance training and
awareness among health care providers, promote a survivor-centered
approach, and create safe spaces for women to disclose violence and
access appropriate care. Additionally, collaborations between health
care institutions, legal authorities, and community organizations are
essential to provide a coordinated and comprehensive response to GBV.

This study contributes to the understanding of the complex nature of GBV
in Afghanistan and provides valuable insights for the development of
targeted interventions and policies. It underscores the importance of
addressing GBV at multiple levels, including individual, interpersonal,
and community-based approaches. Continued research and concerted efforts
are necessary to address the underlying factors contributing to GBV and
to create a society where Afghan women can live free from violence and
oppression.

**References section is created automatically by Rmarkdown. There is no
need to change the references section in the draft file.**

***You shouldn't delete the last 3 lines. Those lines are required for
References section.***

\newpage

# References {#references}

::: {#refs}
:::
